Page not found

### 2023-06-19 manuscript/chapters/1_intro.qmd
all is good :3

![(TODO add phi, add z instead of xi) Every neuron in a rate neural network has a position in the $p$-dimensional embedding space](figures/neural_field_discrete.png)

@@

[^1]: Rigorous proofs of convergence are still an active area of study @CheDua19 and are not the subject of this thesis.

>>>>

![(TODO add phi, add z instead of xi, arrow only in one direction) Every neuron in a rate neural network has a position in the $p$-dimensional embedding space](figures/neural_field_discrete.png)

@@

[^1]: Rigorous proofs of convergence are still an active area of study [@CheDua19;@AGATHENERINE202286] and are not the subject of this thesis.

### 2023-06-19 manuscript/chapters/2_neural_fields.qmd
little kitty be vibing
all is good :3

::: {.content-hidden}
{{< include ../macros.tex >}}
:::

@@

1. Taking Taylor expansions of the activation functions reveals that -- at least up to order 3 -- uneven powers tend to increase stability of the pattern fixed points (and reciprocally, decrease stability of the zero fixed point), and vice-versa for the even powers (with the notable exception of the constant offset, which doesn't play a role in this analysis).
2. A steeper slope at the origin of the activation function also seems to improve stability of the pattern fixed points. This corresponds to a sharper difference between the "inactive" (low potential) and the "active" (high potential) neurons.
3. The spectrum of $K$ is composed of $p$ eigenvalues that depend on the fixed point, and $N-p$ eigenvalues $\lambda_\perp = -1$ corresponding to the orthogonal complement of the pattern fixed points.

@@

![(TODO add spectrum for zero fixed point) Spectrum of $K$ for $N=64$ for the pattern fixed point](figures/spectrum_N=64.png){#fig-spectrum}

@@

$$
\begin{aligned}
h_U(\vec v, t) &= h(\CDF^{-1}(\vec v), t) = h(\CDF^{-1}(v_1), \cdots, \CDF^{-1}(v_p)) \\
w_U(\vec v, \vec u) &= w(\CDF^{-1}(\vec v), \CDF^{-1}(\vec u)).
\end{aligned}
$$

@@

![(TODO) A grid in $[0,1]^2$ and the corresponding samples in $\R^2$. The numeric density of points in $\R^2$ approaches the normal distribution as the grid becomes finer.](figures/sampling_R2_01.png){#fig-sampling-R2-012}

>>>>

::: {.content-hidden}
{{< include ../macros.tex >}}

```{python}
import matplotlib as mpl, matplotlib.pyplot as plt, matplotlib.colors as mplcolors, matplotlib.patheffects as pe, matplotlib.ticker as ticker
import numpy as np
import sys
sys.path.append('../../notebooks')  # import neurodyn
from neurodyn import *
from matplotlib_tufte import *
setup()

mpl.use('svg')

def mktrans(fig, ax):
	"""Makes figure and axes transparent"""
	fig.patch.set_facecolor('#FFFFFF00')
	ax.patch.set_facecolor('#FFFFFF00')
```
:::

@@

1. Taking Taylor expansions of the activation functions reveals that -- at least up to order 3 -- uneven powers tend to increase stability of the pattern fixed points (and reciprocally, decrease stability of the zero fixed point), and vice-versa for the even powers (with the notable exception of the constant offset, which doesn't play a role in this analysis).
2. A steeper slope at the origin of the activation function also seems to improve stability of the pattern fixed points. This corresponds to a sharper difference between the "inactive" (low potential) and the "active" (high potential) neurons.
3. The spectrum of $K$ is composed of $p$ eigenvalues (labeled $\lambda_\mu$ for all $\mu\in\{1,\cdots,p\}$) that depend on the fixed point, and $N-p$ eigenvalues $\lambda_\perp = -1$ corresponding to the orthogonal complement of the pattern fixed points.

@@

```{python}
#| label: fig-spectrum
#| fig-cap: Spectrum of $K$ for $N=64$ and $p=2$. The numerical estimation (black) of the eigenvalues is close to the analytical ($N \to \infty$) derivation (red). Eigenvalues above the dotted gray line(positive values) correspond to unstable fixed points, and below (negative values) correspond to stable fixed points. Only the 
#| fig-subcap:
#|   - Pattern (stable) fixed point
#|   - Zero (unstable) fixed point
#| layout-ncol: 2
#| echo: false

dsigmoid = lambda x: sigmoid(x) * (1 - sigmoid(x))

def plot_spectrum(hstar_str: str):
	params = LowRankRNNParams.new_valentin(p=2, N=256, phi=sigmoid, I_ext=zero, exclude_self_connections=False)
	if hstar_str == 'pattern':
		hstar = params.F[:, 0]
	elif hstar_str == 'zero':
		hstar = np.zeros(params.N)
	dphi = dsigmoid(hstar)
	l = 1/params.N * np.einsum('im,im->m', params.F, params.G * dphi[:, None]) - 1
	l.sort()
	J = np.einsum('im,jm->ij', params.F, params.G * dphi[:, None]) / params.N - np.eye(params.N)

	fig, ax = plt.subplots()
	mktrans(fig, ax)
	ax.axhline(0, color='gray', linestyle='--')
	ax.plot(np.sort(np.linalg.eigvals(J).real)[::-1][:8], marker='o', clip_on=False)
	ax.plot([7,8,8.5], np.sort(np.linalg.eigvals(J).real)[::-1][8:11], linestyle='--', color='k', clip_on=False)
	ax.plot(l[::-1], color='r', marker='o', clip_on=False)
	ax.set_ylim((-1, 0.5))
	ax.set_xlim((0, 8.5))
	ax.set_ylabel('Eigenvalues')
	breathe()
	return fig, ax

# pattern fixed points
fig, ax = plot_spectrum('pattern')
display(fig)

# zero fixed point
fig, ax = plot_spectrum('zero')
display(fig)
```

@@

$$
\begin{aligned}
h_U(\vec v, t) &= h(\CDF^{-1}(\vec v), t) = h(\CDF^{-1}(v_1), \cdots, \CDF^{-1}(v_p)) \\
w_U(\vec v, \vec u) &= w(\CDF^{-1}(\vec v), \CDF^{-1}(\vec u)).
\end{aligned}
$$ {#eq-hu-wu}

@@

![(TODO not color) A grid in $[0,1]^2$ and the corresponding samples in $\R^2$. The numeric density of points in $\R^2$ approaches the normal distribution as the grid becomes finer.](figures/sampling_R2_01.png){#fig-sampling-R2-012}

### 2023-06-19 manuscript/chapters/3_mappings.qmd
all is good :3

For these reasons, we can see why the composition of $w_U$ with a possibly non-differentiable mapping $S$ might be problematic. Since the only thing we know about $S$ is that it is measurable and bijective, *there is no guarantee that the regularity of the kernel is conserved through composition with the mapping $S$.*

@@

@@

[^edges]:
  We note that strictly speaking, $S^n$ maps $(0., 0.)$ onto $0$, and the point
  $$
  S^n(1., 1.) = S^n(0.111\cdots, 0.111\cdots) = 0.111\cdots = 1
  $$
  is never reached at finite $n$. Numerically, we encounter the problem that the inverse of the Gaussian CDF, used to compute the connectivity, goes to infinity when $v_\mu$ (one of the components of $(v_1, v_2) = S^{-1}(\alpha)$) goes to $0$ or $1$. To avoid this, we offset the squares (by $2^{-n-1}$ along each dimension), so that geometrically their position is not indexed by their lower-left corner, but rather their center. Analytically, this is not a problem, because the kernel defined in @eq-hu-wu diverges only linearly, compared to the exponential cutoff of the Gaussian distribution, so that the integrand still vanishes at infinity.

<!-- TODONE NOTE : talk about how we center the bins instead of taking the corners. this is not super important in the $n\to\infty$ limit, but numerically at 0 and 1 the $\mathrm{CDF}^{-1}$ goes to infty. at infinity, the connectivity kernel is infinite but we have zero neurons (zero measure, linear growth of kernel vs exponential tail of the gaussian) : effectively the recurrent current is zero and we can safely ignore these effects. -->

@@

<!-- TODONE NOTE : might need to talk about how $S(v^{(n)}_1,v^{(n)}_2) = S^n(v^{(n)}_1,v^{(n)}_2)$ ? so basically the limit $n\to\infty$ is the limit where we take infinite precision in the input. Or we can drop the $n$ on the inputs, $\alpha^{(n)} = S^n(v^{(n)}_1,v^{(n)}_2) = S^n(v_1,v_2)$ (i think this true just for pointwise convergent mappings) -->

@@

Research in computer science and efficient data structures has given rise to other ways of storing 2-dimensional information in a 1-dimensional memory.
A problem in GPU computing is that of texture locality: sampling textures often involves reading data corresponding to a small 2D region, and therefore to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping" or "Lebesgue curve" @Sagan1994). In this way, points close in 1D *tend* to be tightly packed in 2D. We can define the Z-order curve (Z-mapping) in the following way:[^zcurve]

@@

We also see that the Column mapping is not local in this sense, since there is a large variation along the vertical direction, which seems to always be "of order one". This can be explained by looking at the expression for $C^{-1}_2(\alpha^{(n)}) = 0.b_{n+1} b_{n+2} \cdots b_{2n}$ : small variations of the order of $2^{-n}$ in the value of $\alpha$ always result in variations of order one in $C^{-1}_2$.

@@

The Z-mapping seems to converge to a fractal curve. Numerically, there seems to be an infinite number of discontinuities, but the sizes of these discontinuities are, for the most part, small, so that *on average* they vanish as $n\to\infty$. We won't go further into detail about this,[^zmapping-notbijective] but this gives some intuition for the arguments presented in @sec-locality.

@@

[^zmapping-notbijective]:
  We refer to @Sagan1994 for an in-depth study, and in particular chapter 6, in which the author shows that the Z-curve converges to the Lebesgue curve.<br> 
  <!-- which is a Peano-type function (continuous and surjective from $[0,1]$ into $[0,1]^2$).<br> -->
  Despite what can often be found in online resources, the Z-curve is *not* a bijection.
  <!-- because the added injectivity would violate Netto's theorem. -->
  For every point in $[0,1]^2$, there are at most 4 distinct preimages in $[0,1]$. The reason for this is the ambiguity in the writing of binary expansions, and in @sec-zcurve-noninjective we give proof that the Z-curve is not injective in the $n\to\infty$ limit.<br>
  For the following, we will consider the Z-curve to be bijective by disallowing expansions of the type $0.1000\cdots=0.0111\cdots$.
  <!-- We note that in @Sagan1994 many other Peano curves (continuous surjections $[0,1]\mapsto [0,1]^2$) are presented (albeit more difficult to implement numerically). -->
  <!-- Nicole: I don't say that the Lebesgue curve is continuous, because that opens up a whole other can of worms that I don't want to go into -->

@@

### Visualizating of the mappings of the dynamic 2D neural field

@@

![(TODO) Animation showing the image of the dynamics 2D neural field through different mappings $S : [0,1]^2 \mapsto [0,1]$](figures/photo_2023-06-10_17-21-19.jpg)

>>>>

For these reasons, we can see why the composition of $w_U$ with a possibly non-differentiable mapping $S$ might be problematic. Since the only thing we know about $S$ is that it is measurable and bijective, *there is no guarantee that the regularity (continuity and the differentiability) of the kernel is conserved through composition with the mapping $S$.*

@@

TOOD : merge with previous

@@

[^edges]:
  We note that in our definition, $S^n$ maps $(0., 0.)$ onto $0$. Numerically, we encounter the problem that the inverse of the Gaussian CDF, used to compute the connectivity, goes to infinity when $v_\mu$ (one of the components of $(v_1, v_2) = S^{-1}(\alpha)$) goes to $0$ (or $1$). To avoid this, we offset the squares (by $2^{-n-1}$ along each dimension), so that geometrically their position is not indexed by their lower-left corner, but rather their center. Analytically, this is not a problem, because the kernel defined in @eq-hu-wu diverges only linearly, compared to the exponential cutoff of the Gaussian distribution, so that the integrand still vanishes at infinity.

@@

@@

Research in computer science and efficient data structures has given rise to other ways of storing 2-dimensional information in a 1-dimensional memory.
A problem in GPU computing is that of texture locality: sampling textures often involves reading data corresponding to a small 2D region, and therefore to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping", and similar to the "Lebesgue curve" @Sagan1994). In this way, points close in 1D *tend* to be tightly packed in 2D. We can define the Z-order curve (Z-mapping) in the following way:[^zcurve]

@@

We also see that the Column mapping is not local in this sense, since there is a large variation along the vertical direction, which seems to always be "of order one". This can be explained by looking at the expression for $C^{-1}_2(\alpha^{(n)}) = 0.b_{n+1} b_{n+2} \cdots b_{2n}$ : small variations of the order of $2^{-n}$ in the value of $\alpha$ always result in variations of order one in $C^{-1}_2$.e

@@

The Z-mapping seems to converge to a fractal curve. Numerically, there seems to be an infinite number of discontinuities, but the sizes of these discontinuities are, for the most part, small, so that *on average* they vanish as $n\to\infty$. We won't go further into detail about this, but this gives some intuition for the arguments presented in @sec-locality.

@@

<!--
[^zmapping-notbijective]:
  We refer to @Sagan1994 for an in-depth study, and in particular chapter 6, in which the author shows that the Z-curve converges to the Lebesgue curve.<br> 
  <!-- which is a Peano-type function (continuous and surjective from $[0,1]$ into $[0,1]^2$).<br> 
  Despite what can often be found in online resources, the Z-curve is *not* a bijection.
  <!-- because the added injectivity would violate Netto's theorem. 
  For every point in $[0,1]^2$, there are at most 4 distinct preimages in $[0,1]$. The reason for this is the ambiguity in the writing of binary expansions, and in @sec-zcurve-noninjective we give proof that the Z-curve is not injective in the $n\to\infty$ limit.<br>
  For the following, we will consider the Z-curve to be bijective by disallowing expansions of the type $0.1000\cdots=0.0111\cdots$.
  <!-- We note that in @Sagan1994 many other Peano curves (continuous surjections $[0,1]\mapsto [0,1]^2$) are presented (albeit more difficult to implement numerically). 
  <!-- Nicole: I don't say that the Lebesgue curve is continuous, because that opens up a whole other can of worms that I don't want to go into 
<!-- Nicole: OK, after discussion with valentin, our definition of the Z-mapping is not exactly the same as Lebesgue curve.
  Z-mapping might seem to approximate the Lebesgue curve, but it is not the same.
  Lebesgue is defined on the Cantor set (the cantor set avoids the discontinuities, e.g. around alpha=0.5), then extended linearly, thus continuous
  Our Z-mapping is defined through the bits OBTAINED FROM THE INDICATOR FUNCTIONS; and so it works on NON-AMBIGUOUS binary representations. Thus, our Z-mapping is a bijection.
  Our Z-mapping is also not continuous, consider alpha=0.5. So we don't have a problem with Netto's theorem.
-->

@@

<!-- ### Visualizing of the mappings of the dynamic 2D neural field

@@

![(TODO) Animation showing the image of the dynamics 2D neural field through different mappings $S : [0,1]^2 \mapsto [0,1]$](figures/photo_2023-06-10_17-21-19.jpg) -->

### 2023-06-19 manuscript/chapters/4_locality.qmd
all is good :3

<!-- TODO : we may argue that the Z-curve is not bijective. This complicates the proof. Though, the pre-image of (v_1, v_2) is composed of at most 4 points, and the measurability compensates. -->

@@

<!-- Nicole: V_n -> 0 is implied by continuity -->

@@

![(TODO change title, change name of mappings) (Numerator of the) average binned variation as a function of $n$. The local mappings have a growth slower than the upper bound $2\times 2^n$.](figures/v_num.png){width=500px #fig-vnum}

@@

Let us consider the Z-mapping, and the two terms compositing its inverse. In the case of $\alpha, \alpha^\prime$ above, we have if $n$ is even :

$$
\begin{aligned}
Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3} \cdots i_1}_\text{$n/2$ bits} b_n b_{n+2} \cdots b_{2m-1} \\
Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_0}_\text{$n/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}
\end{aligned}of this locality
$$

@@

<!-- [TODO MATH : prove that it is indeed Lipschitz. The main problem I've encountered with this is that these functions are $\textrm{regular function} \circ \textrm{CDF}^{-1}$, and the inverse CDF goes to infinity at 0 and 1]. -->

>>>>

@@

The definition of locality can essentially be seen as a weakened version of continuity; since from definition the image of a small neighbourhood through a continuous function is also a small neighbourhood. Therefore, continuity of a function $f$ implies that $V_n(f) \xrightarrow{n \to \infty} 0$. The reverse is obviously not true, since for instance the metric $V_n$ vanishes when applied to a step function, which is only piecewise continuous.

@@

![(TODO change title, change name of mappings, remove the theory, but keep upper bound) (Numerator of the) average binned variation as a function of $n$. The local mappings have a growth slower than the upper bound $2\times 2^n$.](figures/v_num.png){width=500px #fig-vnum}

@@

Let us consider the Z-mapping, and the two terms compositing its inverse. In the case of $\alpha, \alpha^\prime$ above, we have if $n$ is even:

$$
\begin{aligned}
Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3} \cdots i_1}_\text{$n/2$ bits} b_n b_{n+2} \cdots b_{2m-1} \\
Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_0}_\text{$n/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}
\end{aligned}
$$

@@
